\emph{Gaia} DR2 identifies VFTS 682 as the most massive runaway star known to date

### 2018-05-14 VFTS682.tex
minor
started shaping HST section. Requires some significant re-writing of Sec. 3 I think

Using the stars selected as described in \Secref{data:gaia} (a
subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
$\langle\mu_\mathrm{RA}\rangle = 1.7240\pm0.0002\,\mathrm{mas\ yr^{-1}}$ and
$\langle\mu_\mathrm{DEC}\rangle = 0.6986\pm0.0003\,\mathrm{mas\
  yr^{-1}}$. These values are
in good agreement with what is found by \cite{lennon:18}. However, we
emphasize their sensitivity to the choice of stars adopted to define
the reference frame. Subtracting these values from the
proper motion of VFTS 682 (see \Tabref{tab:vfts682}), we obtain the
components of proper motion of the star relative to the surrounding region
$\delta\mu_\mathrm{RA} = 0.12\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\delta\mu_\mathrm{DEC} =
0.08\pm 0.08\,\mathrm{mas\ yr^{-1}}$. The error budget is
dominated by the errors on the proper motion components of VFTS 682.

These can be converted into
the components of the relative transverse velocity $\delta v_\mathrm{RA}=28\pm16\,\kms$,
$\delta v_\mathrm{DEC}=20\pm19\,\kms$, assuming a distance of
50\,kpc. % (we do not account for the uncertainty in the distance
% estimate when propagating errors).
The radial velocity from
\cite{bestenlehner:11} then gives the third component along
the line of sight, allowing us to calculate the three-dimensional
peculiar speed of the star:

\begin{equation}
  \label{eq:speed_around}
  v_\mathrm{pec} = \sqrt{\left(\delta v_\mathrm{RA}\right)^2
    +\left(\delta v_\mathrm{DEC}\right)^2+\left(\delta
      v_\mathrm{rad}\right)^2} = 46 \pm 18 
  \kms \ .
\end{equation}
This value for the three-dimensional speed of VFTS 682 with respect the
surrounding stars make it the most massive ``bona fide'' runaway star
known to date.

\subsection{Does it come from the R136 cluster?}
\label{sec:r136_origin}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=0.48\textwidth]{./figures/main_plot_good}  
  \caption{The red solid arrow indicates the proper motion of VFTS 682
    relative to the region, starting from the present day position of
    the star. The orange arrows indicate the possible
    directions of projected motion within the errors, and are extended
    backwards (dashed) to illustrate the uncertainty on the origin of the
    star. The blue stars
    are a subset of those we use to define the local rest
    frame. \todo{load color picture on background}}

@@

\section{Discussion}
\label{sec:discussion}
\todo{subsection on validation with HST data?Discussion on
  correlations between pmRA and pmDEC?Not sure where to start. HST
  data for VFTS 682 dont seem to be public.}
\todo{Selma: Should add caveat that it may be a spurious detection. That reconfirmation in following epochs is warranted. }

>>>>

Using the stars selected as described in \Secref{data:gaia} (a
subset is shown in blue in \Figref{fig:main}), we find averaged proper motion components of
$\langle\mu_\mathrm{RA}\rangle = 1.7240\pm0.0002\,\mathrm{mas\ yr^{-1}}$ and
$\langle\mu_\mathrm{DEC}\rangle = 0.6986\pm0.0003\,\mathrm{mas\
  yr^{-1}}$. These values are
in good agreement with what is found by \cite{lennon:18}. However, we
emphasize their sensitivity to the choice of stars adopted to define
the reference frame. Subtracting these values from the
proper motion of VFTS 682 (see \Tabref{tab:vfts682}), we obtain the
components of proper motion of the star relative to the surrounding region
$\delta\mu_\mathrm{RA}^{Gaia} = 0.12\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\delta\mu_\mathrm{DEC}^{Gaia} =
0.08\pm 0.08\,\mathrm{mas\ yr^{-1}}$. The error budget is
dominated by the errors on the proper motion components of VFTS 682.

These can be converted into
the components of the relative transverse velocity $\delta v_\mathrm{RA}^{Gaia}=28\pm16\,\kms$,
$\delta v_\mathrm{DEC}^{Gaia}=20\pm19\,\kms$, assuming a distance of
50\,kpc. % (we do not account for the uncertainty in the distance
% estimate when propagating errors).
The radial velocity from
\cite{bestenlehner:11} then gives the third component along
the line of sight, allowing us to calculate the three-dimensional
peculiar speed of the star:

\begin{equation}
  \label{eq:speed_around}
  v_\mathrm{pec}^{Gaia} = \sqrt{\left(\delta v_\mathrm{RA}^{Gaia}\right)^2
    +\left(\delta v_\mathrm{DEC}^{Gaia}\right)^2+\left(\delta
      v_\mathrm{rad}^{VFTS}\right)^2} = 46 \pm 18 
  \kms \ .
\end{equation}
This value for the three-dimensional speed of VFTS 682 with respect the
surrounding stars make it the most massive ``bona fide'' runaway star
known to date.

\section{Validation with HST proper motions}

Recently, \citet{platais:18} used HST observations (based on GO-12499
\& 13359 observations with WFC3/UVIS and on archival WFPC2 and
GO-12499 with both cameras, \citealt{platais:15}) to determine the
proper motion of massive stars in the 30 Doradus region. For
large distances ($\gtrsim50$\,kpc), HST data can compete or even
exceed Gaia data in quality.

In the catalog from \citet{platais:18} they listed the proper motion
components of VFTS682 relative to the field, providing a
measure of the motion of this star completely independent from
ours. HST images of VFTS682 are not saturated in any epoch \todo{?},
thanks to the large extinction toward this star. They found proper
motion components relative to the field
$\delta\mu_\mathrm{RA}^\mathrm{HST} = 0.01\pm0.13\,\mathrm{mas\
  yr^{-1}}$ and
$\delta\mu_\mathrm{DEC}^\mathrm{HST}=0.2\pm0.1\,\mathrm{mas\
  yr^{-1}}$.

Since these measures are completely independent from ours, we can take
an average weighted with $1/\delta \mu_i^2$ (cf.~notation in
\Eqref{eq:mean}), obtaining $\delta\mu_\mathrm{RA}=\pm\,\mathrm{mas\
  yr^{-1}}$ and $\delta\mu_\mathrm{DEC}=\pm\,\mathrm{mas\
  yr^{-1}}$.

% %% pure HST physical velocities
% \todo{done in the train: to be checked} Converting these into physical velocities we obtain 
% $\delta v_\mathrm{RA}=2.3\pm30\,\kms$, $\delta v_\mathrm{DEC}=47\pm\,23\kms$, corresponding to
% $v_\parallel^\mathrm{HST} = 47\pm\,1\kms$.

\todo{redo below with average HST+Gaia?Actually redo also calculation
  of $v_\mathrm{pec}$...}
\subsection{Does it come from the R136 cluster?}
\label{sec:r136_origin}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=0.48\textwidth]{./figures/main_plot_good}  
  \caption{The red solid arrow indicates the proper motion of VFTS 682
    relative to the region, starting from the present day position of
    the star. The orange arrows indicate the possible
    directions of projected motion within the errors, and are extended
    backwards (dashed) to illustrate the uncertainty on the origin of the
    star. The blue stars
    are a subset of those we use to define the local rest
    frame. \todo{load color picture on background}\todo{show
      separately arrows fro Gaia/HST and combination? Or show only
      combination? Right now it's Gaia only}}

@@

\section{Discussion}
\label{sec:discussion}
\todo{Selma: Should add caveat that it may be a spurious detection. That reconfirmation in following epochs is warranted. }
